Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks

### 2015-12-07 manuscript/manuscript_wtc3.Rmd
tables #ing are now automated

@@

## Leaf resource distribution
Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`). Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table 1). This was because $\Psi_{pd}$, $\Psi_l$ (Table 1) and E (Table 2) did not differ between leaf types. Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table1). Leaf mass per area was not different between leaf types. No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types.

\
```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/par2.R")
```  
\
**Figure 1**. Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014. Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf. Each date represents the starting date for each measurement campaign. Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.

\
**Table 1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

@@

## Photosynthetic capacity and leaf photosynthesis rates 
The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table 1), as estimated from AC~i~ curves (Figure `r acifig`a). Within leaf types, no effect of the warming treatment was detected on either parameter. Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).

\
```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/aci_vcmaxnitro.R")
```  
\
**Figure 2**. (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light. (c) The relationship between A and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions. For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).

\
Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2). Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c). Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table 2). No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light. Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).

\
**Table 2**. Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa. Different letters represent significant differences between leaf type, light environment and temperature treatments. The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.

@@

## Stomatal conductance and leaf water-use efficiency
On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table 2). Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a). No effect of the warming treatment was detected on g~s~ within or across leaf types. Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).

\
```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```  
\
**Figure 3**. The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are 95% confidence intervals from the mean.

\
```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/physiology_barplots.R")
```  
\
**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.

\
Measured under ambient light and temperature conditions leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001). Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001). Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table 2). The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04. For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).  Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).

\
```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
source("master_scripts/wateruse.R")
```  
\
**Figure 5**. (a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.

\
Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca. 2&permil; (Table 1). No effects of the warming treatment on leaf $\delta$^13^C were detected. Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.

@@

\
```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/delta_cica.R")
```  
\
**Figure 6**. Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light. The solid line represents the theoretical line for C3 plants from Evans et al. [-@evans1986carbon].

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001). Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table 2). Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186). The large increases in g~m~ in shade leaves under high light resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b). No differences in g~m~ were detected with the warming treatment within leaf types. Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively). We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~. Comparison of A-Cc curves (Figure `r accfig`) and A-Ci curves revealed similar differences between sun and shade leaves.

@@

\
```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/cicc_bar.R")
```  
\
**Figure 7**. The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.

>>>>

wtc3tab1 <- "3.1"
wtc3tab2 <- "3.2"

@@

## Leaf resource distribution
Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`). Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`). This was because $\Psi_{pd}$, $\Psi_l$ (Table `r wtc3tab1`) and E (Table `r wtc3tab2`) did not differ between leaf types. Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table1). Leaf mass per area was not different between leaf types. No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types.

\
```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/par2.R")
```  
\
**Figure 3.1**. Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014. Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf. Each date represents the starting date for each measurement campaign. Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.

\
**Table 3.1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

@@

## Photosynthetic capacity and leaf photosynthesis rates 
The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from AC~i~ curves (Figure `r acifig`a). Within leaf types, no effect of the warming treatment was detected on either parameter. Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).

\
```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/aci_vcmaxnitro.R")
```  
\
**Figure 3.2**. (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light. (c) The relationship between A and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions. For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).

\
Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table `r wtc3tab2`). Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c). Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`). No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light. Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).

\
**Table 3.2**. Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa. Different letters represent significant differences between leaf type, light environment and temperature treatments. The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.

@@

## Stomatal conductance and leaf water-use efficiency
On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`). Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a). No effect of the warming treatment was detected on g~s~ within or across leaf types. Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).

\
```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```  
\
**Figure 3.3**. The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are 95% confidence intervals from the mean.

\
```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/physiology_barplots.R")
```  
\
**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.

\
Measured under ambient light and temperature conditions leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001). Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001). Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`). The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04. For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).  Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).

\
```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
source("master_scripts/wateruse.R")
```  
\
**Figure 3.5**. (a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.

\
Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca. 2&permil; (Table `r wtc3tab1`). No effects of the warming treatment on leaf $\delta$^13^C were detected. Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.

@@

\
```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/delta_cica.R")
```  
\
**Figure 3.6**. Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light. The solid line represents the theoretical line for C3 plants from Evans et al. [-@evans1986carbon].

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001). Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table `r wtc3tab2`). Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186). The large increases in g~m~ in shade leaves under high light resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b). No differences in g~m~ were detected with the warming treatment within leaf types. Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively). We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~. Comparison of A-Cc curves (Figure `r accfig`) and A-Ci curves revealed similar differences between sun and shade leaves.

@@

\
```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/cicc_bar.R")
```  
\
**Figure 3.7**. The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.
